Does It Scale? Static Output Approximation of PHP Web Applications Using Symbolic Execution

### 2017-01-11 paper.tex

\documentclass[preprint]{sig-alternate-05-2015}

\setlength\parindent{0pt}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{hyperref}
\usepackage{booktabs}
\usepackage{paralist}
\usepackage{listings,multicol}
\usepackage{lipsum}
\usepackage{cite}
\usepackage{enumerate}
\usepackage{subcaption}
\usepackage{cleveref}
\usepackage{ucs}
\usepackage{framed}

@@

\usepackage[labelfont=bf]{caption}

@@

\lstset{
	language=PHP,
	basicstyle=\ttfamily\tiny,
	basewidth={0.5em,0.4em}, % for other columns settings
	columns=fullflexible,
	numbers=left,
	numberstyle=\tiny,
	frame=single,
	showstringspaces=false,
	xleftmargin=2em,
	framexleftmargin=1.5em,
	backgroundcolor=\color{lightgrey}
}

@@

@@

% TIMES
\usepackage{mathptmx}% http://ctan.org/pkg/mathptmx

@@

\usepackage{xcolor}
%\usepackage[dvipsnames]{xcolor}

\definecolor{lightgrey}{HTML}{EEEEEE}

\definecolor{orange}{RGB}{230, 100, 20}
\definecolor{gruen}{RGB}{255, 190, 69}

\definecolor{RYB1}{HTML}{DCDCDC}
\definecolor{RYB2}{HTML}{C0C0C0}
\definecolor{RYB3}{HTML}{808080}

\definecolor{RYB4}{HTML}{DCDCDC}
\definecolor{RYB5}{HTML}{C0C0C0}
\definecolor{RYB6}{HTML}{808080}

\definecolor{RYB7}{HTML}{DCDCDC}
\definecolor{RYB8}{HTML}{C0C0C0}
\definecolor{RYB9}{HTML}{808080}

\definecolor{RYB10}{HTML}{D3D3D3}
\definecolor{RYB11}{HTML}{A9A9A9}

\pgfplotscreateplotcyclelist{ColorListBar}{
{fill=RYB1},
{fill=RYB2},
{fill=RYB3},
}

\pgfplotscreateplotcyclelist{ColorListBar2}{
{fill=RYB4},
{fill=RYB5},
{fill=RYB6},
}

\pgfplotscreateplotcyclelist{ColorListBar3}{
{fill=RYB7},
{fill=RYB8},
{fill=RYB9},
}

\begin{document}
\lstset{
	language=PHP,
	basicstyle=\footnotesize\ttfamily,
    commentstyle = \color{gray},
    extendedchars = \true,
    inputencoding = utf8x,
    keepspaces = true,
    keywordstyle = \bfseries,
    frame = tb
}

% Copyright
\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
%\setcopyright{rightsretained}
%\setcopyright{usgov}
%\setcopyright{usgovmixed}
%\setcopyright{cagov}
%\setcopyright{cagovmixed}

% DOI
\doi{10.475/123_4}

% ISBN
\isbn{123-4567-24-567/08/06}

%Conference
\conferenceinfo{ISSTA '17}{July 9--13, 2013, Santa Barbara, CA, USA}

\acmPrice{\$15.00}

%
% --- Author Metadata here ---
\conferenceinfo{ISSTA '17}{July 9--13, 2013, Santa Barbara, CA, USA}
%\CopyrightYear{2007} % Allows default copyright year (20XX) to be over-ridden - IF NEED BE.
%\crdata{0-12345-67-8/90/01}  % Allows default copyright data
% (0-89791-88-6/97/05) to be over-ridden - IF NEED BE.
% --- End of Author Metadata ---

\title{Does It Scale? Static Output Approximation of PHP Web Applications Using
Symbolic Execution}

%

\numberofauthors{3} %  in this sample file, there are a *total*
% of EIGHT authors. SIX appear on the 'first-page' (for formatting
% reasons) and the remaining two appear in the \additionalauthors section.
%
\author{
% You can go ahead and credit any number of authors here,
% e.g. one 'row of three' or two rows (consisting of one row of three
% and a second row of one, two or three).
%
% The command \alignauthor (no curly braces needed) should
% precede each author name, affiliation/snail-mail address and
% e-mail address. Additionally, tag each line of
% affiliation/address with \affaddr, and tag the
% e-mail address with \email.
%
% 1st. author
\alignauthor
Stefan Mühlbauer\\
       \affaddr{TU Braunschweig, Germany}\\
       %\email{s.muehlbauer@tu-bs.de}
% 2nd. author
\alignauthor
Christian Kästner\\
      \affaddr{Carnegie Mellon University, USA}\\
       %\email{ckaestner@cs.cmu.edu}
% 3rd. author
\alignauthor
Tien N. Nguyen\\
       \affaddr{University of Texas, USA}\\
%      %\email{tien.n.nguyen@utdallas.edu}
}
% There's nothing stopping you putting the seventh, eighth, etc.
% author on the opening page (as the 'third row') but we ask,
% for aesthetic reasons that you place these 'additional authors'
% in the \additional authors block, viz.
\additionalauthors{Additional authors: John Smith (The Th{\o}rv{\"a}ld Group,
email: {\texttt{jsmith@affiliation.org}}) and Julius P.~Kumquat
(The Kumquat Consortium, email: {\texttt{jpkumquat@consortium.net}}).}
\date{30 July 1999}
% Just remember to make sure that the TOTAL number of authors
% is the number that will appear on the first page PLUS the
% number that will appear in the \additionalauthors section.

\maketitle
\begin{abstract}
Dynamic web applications have become widely popular and are to
a large proportion based on the scripting language PHP. Output approximation of web
applications enables a range of additional tool support as well as possibilities
for vulnerability detection. Unfortunately, recent approximation approaches have
only been evaluated for smaller systems.

@@

%
%  Use this command to print the description
%
\printccsdesc

@@

\keywords{Static program behavior, Static metrics, Dynamic language features,
Static Analysis, PHP}

\section{Introduction}
With the emerging world wide web, dynamic web applications have become
widely popular. Various different implementation techniques have emerged,
ranging from technologies for programming languages (JSP, ASP .NET) and web
application frameworks for  script languages (Ruby On Rails, Django) to
languages tailored specifically to the domain of web applications. PHP \cite{phpNET} is
a programming language focused on server-side application development. As of 2012,
it was used by 78.8 percent of the ten million most popular websites (according
to Alexa popularity ranking) \cite{alexaPHP}, ranking 7th on the TIOBE
programming community index \cite{tiobePHP}, and was the ranked as the 6th most
popular language on GitHub \cite{githubPHP}.

@@

%Symbolic execution extends concrete execution semantics with semantics to
%handle symbolic values. Hence, concrete execution semantics is contained in
%symbolic execution semantics. Symbolic execution has been used and studied for generating test suites, but
%has more recently also been applied for approximating the output of web
%applications.

@@

\begin{compactitem}
	\item \emph{PhpSync}: Using all variants of  HTML client page output, every
	single one can be statically checked for Markup Validity, i.e., if the web page
	conforms to syntactical specifications for HTML and other client-side
	languages. After validation errors are detected, auto-fixes are provided.
	Otherwise, the validation error can be traced back to source code responsible
	for the defect, which then can be refactored manually
	\cite{Nguyen:2011:AFH:2190078.2190142}.

@@

	\item \emph{Varis}: Editor services such as “jump to declaration” are
	nontrivial for dynamic web applications due to their staged nature. Varis
	provides editor services across stages for client-side code: Starting with an
	ouput approximation a callgraph with conditional edges is constructed, which
	allows navigation in client-site code (HTML, JS and CSS) although it is
	embedded in server-side code
	\cite{Nguyen:2015:VIS:2819009.2819140,Nguyen:2014:BCG:2635868.2635928}.
\end{compactitem}

@@

\subsection{What Went Wrong? Experience Report}

@@

%
\begin{compactitem}
	\item four small-scale systems that selected from the previously mentioned case
	study in order to compare our results with the previous symbolic execution
	engine \cite{Nguyen:2014:BCG:2635868.2635928},

@@

	\item four small-scale systems selected from a list of recent Content
	Management Systems \cite{codegeekz}.
\end{compactitem}

@@

	\end{subfigure}
	\caption{Results}
\end{figure}
\section{Lessons Learned}

\section{Related Work}
\begin{itemize}
	\item Hills et al. \cite{Hills:2013:ESP:2483760.2483786}
	\item Output approximation using context-free grammar construction; applications \cite{minamide2005static,wassermann2007sound}
\end{itemize}

%References
\bibliographystyle{abbrv}
\bibliography{bibliography}

>>>>

\documentclass[sigconf]{acmart}

\setlength\parindent{0pt}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{hyperref}
\usepackage{blindtext}
\usepackage{booktabs}
\usepackage{paralist}
\usepackage{listings,multicol}
\usepackage{lipsum}
%\usepackage{cite}
\usepackage{enumerate}
\usepackage{subcaption}
\usepackage{cleveref}
\usepackage{ucs}
\usepackage{framed}

\renewcommand{\sf}[1]{\textsf{#1}}
\renewcommand{\tt}[1]{\texttt{#1}}

@@

%\usepackage[labelfont=bf]{caption}

\definecolor{lightgrey}{HTML}{FFFFFF}

@@

\lstset{
	language=PHP,
	extendedchars = \true,
	basicstyle=\ttfamily\footnotesize,
	basewidth={0.5em,0.4em}, % for other columns settings
	commentstyle = \color{gray},
	columns=fullflexible,
	numbers=left,
	numberstyle=\tiny,
	frame=tb,
	showstringspaces=false,
	xleftmargin=2em,
	framexleftmargin=1.5em,
	backgroundcolor=\color{lightgrey},
	inputencoding = utf8x,
    keepspaces = true,
    keywordstyle = \bfseries,
}

@@

\newcommand*{\origrightarrow}{}
\let\oldarrow\textrightarrow
\renewcommand*{\textrightarrow}{\fontfamily{cmr}\selectfont\origrightarrow}

@@

% TIMES
%\usepackage{mathptmx}% http://ctan.org/pkg/mathptmx

@@

\usepackage{xcolor}

% Copyright
%\setcopyright{none}
%\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
\setcopyright{rightsretained}
%\setcopyright{usgov}
%\setcopyright{usgovmixed}
%\setcopyright{cagov}
%\setcopyright{cagovmixed}

% DOI
\acmDOI{10.475/123_4}

% ISBN
\acmISBN{123-4567-24-567/08/06}

%Conference
\acmConference[ISSTA'17]{ACM Woodstock conference}{July 2017}{Santa Barbara,
California, USA}
\acmYear{1997}
\copyrightyear{2016}

\acmPrice{15.00}

\begin{document}
\title{Does It Scale? Static Output Approximation of PHP Web Applications Using
Symbolic Execution}
\titlenote{Produces the permission block, and copyright information}
%\subtitle{Extended Abstract}
%\subtitlenote{The full version of the author's guide is available as  \texttt{acmart.pdf} document}

\author{Stefan Mühlbauer}
\affiliation{%
  \institution{Technische Universität Braunschweig}
  %\streetaddress{P.O. Box 1212}
  %\city{Braunschweig} 
  %\state{Germany} 
  %\postcode{43017-6221}
}
%\email{s.muehlbauer@tu-bs.de}

\author{Christian Kästner}
%\orcid{1234-5678-9012}
\affiliation{%
  \institution{Carnegie Mellon University}
  %\streetaddress{5000 Forbes Avenue}
  %\city{Pittsburgh} 
  %\state{PA} 
  %\postcode{15217}
}
%\email{s.muehlbauer@tu-bs.de}

\author{Tien N. Nguyen}
%\orcid{1234-5678-9012}
\affiliation{%
  \institution{University of Texas at Dallas}
  %\streetaddress{P.O. Box 1212}
  %\city{Braunschweig} 
  %\state{Germany} 
  %\postcode{43017-6221}
}
%\email{s.muehlbauer@tu-bs.de}

% The default list of authors is too long for headers}
\renewcommand{\shortauthors}{S. Mühlbauer et al.}

\begin{abstract}
Dynamic web applications have become widely popular and are to
a large proportion based on the scripting language PHP. Output approximation of web
applications enables a range of additional tool support as well as possibilities
for vulnerability detection. Unfortunately, recent approximation approaches have
only been evaluated for smaller systems.

@@

@@

\keywords{ACM proceedings, \LaTeX, text tagging}

\maketitle

\section{Introduction}
With the emerging world wide web, dynamic web applications have become
widely popular. Various different implementation techniques have emerged,
ranging from technologies for programming languages (JSP, ASP .NET) and web
application frameworks for  script languages (Ruby On Rails, Django) to
languages tailored specifically to the domain of web applications. PHP
cite{phpNET} is a programming language focused on server-side application
development. As of 2012, it was used by 78.8 percent of the ten million most popular websites (according
to Alexa popularity ranking) \cite{alexaPHP}, ranking 7th on the TIOBE
programming community index \cite{tiobePHP}, and was the ranked as the 6th
most popular language on GitHub \cite{githubPHP}.

@@

@@

\begin{itemize}
	\item {\emph{PhpSync}: Using all variants of  HTML client page output, every
	single one can be statically checked for Markup Validity, i.e., if the web page
	conforms to syntactical specifications for HTML and other client-side
	languages. After validation errors are detected, auto-fixes are provided.
	Otherwise, the validation error can be traced back to source code responsible
	for the defect, which then can be refactored manually
	\cite{Nguyen:2011:AFH:2190078.2190142}. }

@@

	\item \emph{Varis}: Editor services such as “jump to declaration” are
	nontrivial for dynamic web applications due to their staged nature. Varis
	provides editor services across stages for client-side code: Starting with an
	ouput approximation a callgraph with conditional edges is constructed, which
	allows navigation in client-site code (HTML, JS and CSS) although it is
	embedded in server-side code
	\cite{Nguyen:2015:VIS:2819009.2819140,Nguyen:2014:BCG:2635868.2635928}.
\end{itemize}

@@

\subsection{Things We Learned The Hard Way}
\textcolor{blue}{\blindtext}

\subsubsection{Symbolic vs Concrete Execution}
\textcolor{blue}{\blindtext}

\subsubsection{Multi-Target Method Calls}
The first non-conceptual problem we encountered, in particular for
\sf{WordPress} is related to object-oriented programming. While exploring
different execution paths when the control flow is split, a variable can have
different values depending on the path condition under which a value has been
assigned to it (see the code example in Figure \ref{fig:output_example}). Similar, a variable
can point to different objects depending on the path condition as the snippet in
Figure \ref{fig:multi_target} illustrates. The variable \tt{\$person} can either
point to object \tt{Alice} or \tt{Bob}. In addition, for any method invocated on
that variable, the actual target is ambiguous as there are two, depending on the
path condition. In order to have a sound symbolic execution, both method calls
need to be executed. Nevertheless, the number of contexts for a method can grow
rapidly, and easily become infeasible to execute.

We implemented several heuristic approaches to tame this state space explosion
for multi-target method calls, including executing a method only for a subset or
just a single object. This trade-off between accuracy and soundness
of the symbolic execution and performance time required.

\subsubsection{Dynamic Language Features}
Conceptually, symbolic execution failed for two types of language features. As
stated in Section \ref{sec:workflow}, dynamic web applications are staged and client page
output is computed at runtime. Symbolic execution hand has shown to be a useful
technique to approximate dynamically generated output, it is challenged by
language features that require concrete information rather than symbolic
information at runtime. In particular, this applies to language features that
provide access to reusable code of the web application: include expressions and
function calls.

First, include expressions in PHP usually require a string value representing
the path to the script file or template to include. If this string value is
statically provided by a string literal or can be constructed unambiguously, the
symbolic execution engine was able to resolve the include expression and include
the desired file. However, it is also common to assemble include expressions not
only static string literals, but from several sources including user input
information, database query results or configuration files. Figure
\ref{fig:dynamic_includes} shows a small example of a dynamic include containing
information retrieved from a database. Since any input for a symbolically
executed application is symbolic this information leads to include expressions
containing symbolic values. Also, configuration files represent only one
particular configuration or even just default values or placeholders. Hence, 
symbolic execuion is prone to fail for dynamic includes if symbolic information
is contained.

Second, function calls can either be direct by statically providing the name of
the function (or method), or indirect. Indirect function calls in PHP are
enabled by using built-in functions like \tt{call\_user\_func},
\tt{call\_user\_func\_array} or simply passing a string value with a direct
function call to the \tt{eval} function. The first two functions take as
arguments the name of the function two call, and all arguments to that functions either as additional arguments or
array respectively. The eval function evaluates a given PHP expression passed
as a string literal.
Again, we encountered similar issues to dynamic include expressions with
indirect function call mechanisms. The example in Figure
\ref{fig:indirect_calls} illustrates two functions \tt{header\_serif} and
\tt{header\_sserif} which both print a HTML headline. Depending on whether the
value of the variable \tt{\$style} retrieved from a database is \tt{serif} or
\tt{sserif}, a different function is called. Given that the function name is 
concrete at runtime, the desired function can be called. Otherwise, if the
function name is assembled at runtime and contains symbolic values, no function
can be called as the name is ambiguous or unknown. For this example, symbolic
execution is not able to determine which function to call indirectly.

% code examples for 3.3 experience report
\begin{figure}[t]
	\begin{subfigure}[b]{0.48\textwidth}

\begin{lstlisting}
<?php
class Person {
   function __construct__($name) {
      $this->name = $name;
   }
   function greet($greeting) {
      echo $greeting . $this->name;
   }
}

if (...) {
   $person = new Person("Alice");
} else {
   $person = new Person("Bob");
}
$person->greet("Good Morning, ");
\end{lstlisting}
	\caption{
		Dynamic dispatch: Multiple targets for method calls
		\label{fig:multi_target}
	}
	\end{subfigure}
	\vspace{2mm}
	\hfill
	\begin{subfigure}[b]{0.48\textwidth}

\begin{lstlisting}
<?php
define("ROOT", getcwd());
define("TEMPLATES", ROOT . "templates/")

$template = mysql_result(...);

require_once TEMPLATE . $template;
\end{lstlisting}
	\caption{
		Dynamic Include Resolution
		\label{fig:dynamic_includes}
	}
	\end{subfigure}
	\vspace{2mm}
	\hfill
	\begin{subfigure}[b]{0.48\textwidth}

\begin{lstlisting}
<?php
function header_serif($title) {
   $style = "font-family: serif;";
   echo "<h1 style = '$style'>$title</h1>";
}
function header_sserif($title) {
   $style = "font-family: sans-serif;";
   echo "<h1 style = '$style'>$title</h1>";
}
\end{lstlisting}
	\caption{
		Function calls by indirect invocation mechanisms
		\label{fig:indirect_calls}
	}

	\end{subfigure}
	\caption{
		Code Examples of defective code features
		\label{fig:defective_code_features}
	}
\end{figure}

\subsection{Things We Tried\ldots}
\textcolor{blue}{\blindtext\blindtext}

@@

%
\begin{itemize}
	\item four small-scale systems that selected from the previously mentioned case
	study in order to compare our results with the previous symbolic execution
	engine \cite{Nguyen:2014:BCG:2635868.2635928},

@@

	\item four small-scale systems selected from a list of recent Content
	Management Systems \cite{codegeekz}.
\end{itemize}

@@

	\end{subfigure}
	\caption{Results}
\end{figure}

\section{Lessons Learned}
\textcolor{blue}{\Blindtext}

\section{Related Work}
\textcolor{blue}{\blindtext\blindtext}

%References
\bibliographystyle{ACM-Reference-Format}
\bibliography{bibliography}
